---
title: "Robustness Check: 3 lags"
author: "Camilla Zhang"
date: '2022-10-24'
output: html_document
---
### Import Libraries
```{r}
library(dplyr)
library(countrycode)
library(stringr)
library(plm)
library(tidyr)
library(data.table)
library(ggplot2)
library(Hmisc)
library(zoo)
library(ggplot2)
library(ggthemes)
library(purrr)
library(rworldmap)
library(psych)
```
### Import datasets
```{r}
desta = read.csv("/Users/jayzee/Documents/qmss_fall22/thesis//data/desta.csv")
withdrawals = read.csv("/Users/jayzee/Documents/qmss_fall22/thesis/data/withdrawals.csv")
withdrawals = withdrawals[!duplicated(withdrawals),]  #remove duplicates for withdrawals
desta_dyads = read.csv("/Users/jayzee/Documents/qmss_fall22/thesis/data/desta_dyads_v3.csv")
desta_dyads1 = desta_dyads[!duplicated(desta_dyads),] #Remove duplicate entries in dyadic data
hdi = read.csv("/Users/jayzee/Documents/qmss_fall22/thesis/data/hdi_composite.csv", row.names=NULL)
vdem = read.csv("/Users/jayzee/Documents/qmss_fall22/thesis/data/vdem_folder/vdem.csv", row.names=NULL)
gallup = read.csv("/Users/jayzee/Documents/qmss_fall22/thesis/data/gallup.csv")
```
### Test subset data (to ensure wrangling merging was successful)
```{r}
#desta_dyads1 = subset(desta_dyads, base_treaty %in% c(1, 192, 65, 810, 562))
#desta_dyads1 = subset(desta_dyads, base_treaty %in% c(192, 1, 249))
```

### Divide dataset between base, accession, and consolidated
```{r}
#Base 
base_ptas = subset(desta_dyads1, entry_type == "base_treaty")

#Consolidated
consolidated_ptas = subset(desta_dyads1, entry_type == "consolidated")

#Accession 
accession_ptas = subset(desta_dyads1, entry_type == "accession")
```


###Merge DESTA data with dyadic data to aggregate by country-PTA-year
```{r}
#Join DESTA indices with BASE PTAs based on base treaty number and year
merged_base = merge(base_ptas, desta[,c("base_treaty", "year", "depth_index", "depth_rasch")], by = c("base_treaty", "year"), all.x = TRUE)

#Join DESTA indices with BASE PTAs based on *new* treaty number and year for consolidate PTAs that changed their provisions/depth (The consol. PTAs with updated depth index will get an index, others an NA)
merged_con = merge(consolidated_ptas, desta[,c("number", "depth_index", "depth_rasch")], by = c("number"), all.x = TRUE)
#Subset consolidated data where depth is not NA
con_non_nas = merged_con[!is.na(merged_con$depth_index),]
#Bind non NA consolidated data w/merged data
con_base = rbind(con_non_nas, merged_base)

#Subset CONSOLIDATED PTAs with NA and populate it (These are NAs b/c their depth/provisions didn't change)
con_nas <- merged_con[is.na(merged_con$depth_index),] #Subset
con_nas2 = merge(con_nas, desta[,c("base_treaty", "depth_index")], by = c("base_treaty"), all.x = TRUE) %>% #populate based on base treaty
dplyr::rename(depth_index = depth_index.y)  
merged_con2 = subset(con_nas2, select = -c(depth_index.x)) #remove duplicate columns

# Join the CON BASE and (previously NA) CONSOLIDATED datasets together
con_base2 = rbind(merged_con2, con_base)
con_base2 <- con_base2[!is.na(con_base2$depth_index),]

# Remove duplicate rows based on country, base treaty, year, and depth index
con_base3 = con_base2[!duplicated(con_base2[c("iso1", "depth_index", "base_treaty", "year")]),] 

# Rename depth index column to depth_filled
con_base3 = con_base3 %>% dplyr::rename(depth_filled = depth_index)
```

### Appending accession countries 
```{r}
#Add missing years by base treaty and fill PTA count with 0
desta2 = desta %>% group_by(base_treaty) %>% tidyr::complete(year = min(year):2021, fill = list(depth = NA))

#Populate the depth based on previous year's depth index
desta2$depth_filled <- na.locf(desta2$depth_index) 

#Create accession variable for the new DESTA
merged_acc = merge(accession_ptas, desta2[,c("depth_filled", "depth_rasch", "base_treaty", "year")], by = c("base_treaty", "year"), all.x = TRUE)

#Bind accession with con/base PTAs
merged_all = rbind(con_base3, merged_acc)

#Rename ISO column
merged_all = merged_all %>% dplyr::rename(iso3n = iso1)
```


### Merge DESTA data with GWP incumbent support data
```{r}
#Merge by iso3 numeric country codes
gallup$iso3n = countryname(gallup$Geography, destination = 'iso3n')

#Merge by year
gallup = gallup %>% 
  dplyr::rename(year = Time)

#Remove NAs where country and approval ratings are NA 
gallup = subset(gallup, iso3n != "NA") 
gallup = subset(gallup, Approve != "NA")

#Remove percentages in approval rating variable
gallup$Approve <- as.integer(gsub('%', '', gallup$Approve))

merged_all = merge(merged_all, gallup, by = c("iso3n", "year"), all.x = TRUE)
```
# Find the mean and median of treaties
```{r}
#Get average difference between year and entryforceyear
merged_all$difference_year = merged_all$entryforceyear-merged_all$year
mean(merged_all$difference_year, na.rm = T)
median(merged_all$difference_year, na.rm = T)
#Percentage of trade agreements with a depth index
sum(unique(desta$base_treaty))/sum(unique(desta_dyads1$base_treaty))

```

###Merge PTA data w/HDI + Add a binary indicator to signify if trade agreement is deep
```{r}
## Merge HDI with desta data ##

#Subset HDI dataset
hdi = hdi %>% dplyr::select(iso3, country, region, hdi_1990, hdi_1991, hdi_1992, hdi_1993, hdi_1994, hdi_1995, hdi_1996, hdi_1997, hdi_1998, hdi_1999, hdi_2000, hdi_2001, hdi_2002, hdi_2003, hdi_2004, hdi_2005, hdi_2006, hdi_2007, hdi_2008, hdi_2009, hdi_2010, hdi_2011, hdi_2012, hdi_2013, hdi_2014, hdi_2015, hdi_2016, hdi_2017, hdi_2018, hdi_2019, hdi_2020, hdi_2021, gnipc_2005, gnipc_2006, gnipc_2007, gnipc_2008, gnipc_2009, gnipc_2010, gnipc_2011, gnipc_2012, gnipc_2013, gnipc_2014, gnipc_2015, gnipc_2016, gnipc_2017, gnipc_2018, gnipc_2019, gnipc_2020, gnipc_2021, mys_2005, mys_2006, mys_2007, mys_2008, mys_2009, mys_2010, mys_2011, mys_2012, mys_2013, mys_2014, mys_2015, mys_2016, mys_2017, mys_2018, mys_2019, mys_2020, mys_2021)

#Produce iso3 numeric country codes for hdi_hdi dataset
hdi$iso3n = countryname(hdi$country, destination = 'iso3n')

#Merge HDI with PTA data
hdi_hdi = hdi %>%
  tidyr::pivot_longer(
    cols = starts_with("hdi_"),
    names_to = "year",
    names_prefix = "hdi_",
    values_to = "hdi_score"
  )

hdi_hdi = hdi_hdi %>% select(iso3n, year, hdi_score)
depth_data = merge(merged_all, hdi_hdi, by = c("iso3n", "year"), all.x = T) %>% mutate(developing = case_when(hdi_score < 0.8 ~ 1,
                                           hdi_score>= 0.8 ~ 0))

# Assign NAs in developing to a 1 or 0 based on current values
depth_data$developing2 = nafill(depth_data$developing, type = "nocb")
depth_data$developing3 = ifelse(depth_data$developing2 == 0, "Developed", "Developing")

#Signify that an agreement is deep w/binary variable
depth_data$deep = ifelse(depth_data$depth_filled >= mean(depth_data$depth_filled, na.rm = T), 1, 0)
```


#Descriptive Statistics and EDA

### Data Visualization 1: PTAs signed and Average PTA depth over time b/w developed and developing
```{r}
#Count number of PTAs for each country_type-year (developing/not developing)
country_yr_n = depth_data %>% group_by(developing3,year)  %>% summarise(n = n())
country_yr_n = subset(country_yr_n, !is.na(developing3)) #Remove NA rows
country_yr_n = subset(country_yr_n, !is.na(n))
country_yr_n

#Get average depth of PTAs for each country-year
country_yr_depth = depth_data %>% group_by(developing3,year) %>% summarise(avg_depth = mean(depth_filled)) 
country_yr_depth = subset(country_yr_depth, !is.na(developing3)) #Remove NA rows
country_yr_depth = subset(country_yr_depth, !is.na(avg_depth))
country_yr_depth

country_n = depth_data %>% group_by(country1) %>% summarise(n = n())
country_depth = depth_data %>% group_by(country1) %>% summarise(avg_depth = mean(depth_filled))
country_n

#Add in developing country column 
country_depth_m = merge(country_depth, depth_data[,c('developing3', 'country1')], by = 'country1', all.x = TRUE)
country_depth_m = subset(country_depth_m[!duplicated(country_depth_m), ], !is.na(avg_depth))
country_depth_m

country_n_m = merge(country_n, depth_data[,c('developing3', 'country1')], by = 'country1', all.x = TRUE)
country_n_m = subset(country_n_m[!duplicated(country_n_m), ], !is.na(n))
country_n_m


#Distribution of average depth index by developing-year
depth_dy = ggplot(country_yr_depth, aes(x = year, y = avg_depth, color = as.factor(developing3), group = as.factor(developing3))) +
  geom_line()+
  geom_point()+
  scale_color_manual(values=c('Black','Dark Grey'))+
  ylab("Depth")+
  xlab("Year")+
  labs(color = "") +
  scale_fill_discrete(labels=c('Developed', 'Developing')) +
  theme_tufte()

# Distribution of average count by developing-year
count_dy = ggplot(country_yr_n, aes(x = year, y = n, color = as.factor(developing3), group = as.factor(developing3))) +
  geom_line()+
  geom_point()+
  scale_color_manual(values=c('Black','Dark Grey'))+
  ylab("Count")+
  xlab("Year")+
  labs(color = "") +
  scale_fill_discrete(labels=c('Developed', 'Developing')) +
  theme_tufte()

count_dy
depth_dy
```

```{r}
# Get distribution of depth index
 depth_distrib = ggplot(desta, aes(x = depth_index)) + 
  geom_histogram()+
  labs(title="Distribution of trade agreement depth", x = "Depth Index", y =  "")+
  theme_tufte()

# Get distribution of depth index (sqrt transformation)
depth_distrib_sqrted = ggplot(desta, aes(x = sqrt(depth_index))) + 
  geom_histogram()+
  labs(title="Distribution of trade agreement depth", x = "Depth Index", y =  "")+
  theme_tufte()


depth_distrib
depth_distrib_sqrted
```

#Add 3 control variables

###1. Adding GDP 
```{r}
#Merge GDP with pta data
hdi_gdp = hdi %>%
  tidyr::pivot_longer(
    cols = starts_with("gnipc_"),
    names_to = "year",
    names_prefix = "gnipc_",
    values_to = "gdp"
  )

hdi_gdp = hdi_gdp %>% select(iso3n, year, gdp)
depth_data = merge(depth_data, hdi_gdp, by = c("iso3n", "year"), all.x = T)

# Get distribution of GDP 
gdp_distrib = ggplot(depth_data, aes(x = log(gdp))) + 
  geom_histogram()+
  labs(title="Distribution of GDP", x = "GDP", y =  "")+
  theme_tufte()
gdp_distrib
```

###2. Adding mean yrs of schooling (EDU)
```{r}
#Merge school with PTA data
hdi_edu = hdi %>%
  tidyr::pivot_longer(
    cols = starts_with("mys_"),
    names_to = "year",
    names_prefix = "mys_",
    values_to = "edu"
  )

hdi_edu = hdi_edu %>% select(iso3n, year, edu)
depth_data = merge(depth_data, hdi_edu, by = c("iso3n", "year"), all.x = T)

edu_distrib = ggplot(depth_data, aes(x = edu)) + 
  geom_histogram()+
  labs(title="Distribution of education", x = "education (in mean years of schooling)", y =  "")+
  theme_tufte()
edu_distrib
```

###3. Adding Regime Type
```{r}
#Convert vdem country names to iso3n
vdem = vdem[,c("country_name", "country_text_id", "country_id", "year", "v2x_regime")]
vdem$iso3n = countrycode(vdem$country_text_id, origin = 'iso3c', destination = 'iso3n')
#merge regime with pta data
depth_data = merge(depth_data, vdem[,c("iso3n", "v2x_regime", "year")], by = c("iso3n", "year"), all.x = TRUE)

regime_distrib = ggplot(depth_data, aes(x = v2x_regime)) + 
  geom_histogram()+
  labs(title="Distribution of regime type", x = "Regime score", y =  "")+
  theme_tufte()
regime_distrib
```

# Get unique countries and subset to time period
```{r}
#Subset data down to 2005-2021 and non-NA values
depth_data2 = depth_data[depth_data$year > 2004,]
depth_data2 = depth_data2[!is.na(depth_data2$Approve),]
length(unique(depth_data2$iso3n))
length(unique(depth_data2[depth_data2$developing == 1,]$iso3n))
```

# Models

## PTA Depth Level Analysis (w/year as independent variable)

### Model 3: Depth w/Lags and logged Depth/GDP
```{r}
fe_mod1 <- plm(Approve ~ Lag(sqrt(depth_filled), shift = 3) +
                  developing +
                  Lag(log(gdp)) +
                  Lag(edu) +
                  Lag(v2x_regime) +
                  year,
                    data = subset(depth_data2),
                    index = c("iso3n", "year"), 
                    model = "within",
                    effect = "twoways")
summary(fe_mod1)
```



## Country-Count Level Analysis (w/year as independent variable)
```{r}
#Count all PTAs for each country-year (to get change)
desta_dyads1 = desta_dyads1 %>% rename(iso3n = iso1)
pta_count = desta_dyads1 %>% dplyr::group_by(iso3n, year) %>% dplyr::summarise(count_pta = n())

#Count all PTA withdrawals for each country-year
withdrawals = withdrawals %>% rename(iso3n = iso1)
withdrawals_count = withdrawals %>% dplyr::group_by(iso3n, year) %>% dplyr::summarise(count_withdrawals = n())

#Merge withdrawals column to desta_country_yr
count_data = merge(pta_count, withdrawals_count, by = c("year", "iso3n"), all.x = TRUE, all.y = TRUE)

#Add missing years
count_data <- setDT(count_data)[CJ(iso3n=iso3n, year=seq(min(year), 2021), unique=TRUE),
              on=.(iso3n, year)]

#fill PTA count with 0
count_data <- count_data %>% mutate(count_pta = ifelse(is.na(count_pta), 0, count_pta),
                                                  count_withdrawals = ifelse(is.na(count_withdrawals), 0, count_withdrawals))

#Add the PTAs
count_data = count_data %>% group_by(iso3n) %>%
    mutate(across(count_pta, ~ accumulate(., `+`)))
count_data = count_data %>% group_by(iso3n) %>%
    mutate(across(count_withdrawals, ~ accumulate(., `+`)))

#Create total PTA column
count_data$total_ptas = count_data$count_pta - count_data$count_withdrawals

## Merge approval ratings w/data ##
count_data = merge(count_data, gallup, by = c("iso3n", "year"), all.x = TRUE)

## Merge HDI w/data ##
count_data = merge(count_data, hdi_hdi, by = c("iso3n", "year"), all.x = T) %>% mutate(developing = case_when(hdi_score < 0.8 ~ 1,
                                           hdi_score>= 0.8 ~ 0))

## Merge Control variables w/data ##
count_data = merge(count_data, hdi_gdp, by = c("iso3n", "year"), all.x = T)
count_data = merge(count_data, hdi_edu, by = c("iso3n", "year"), all.x = T)
count_data = merge(count_data, vdem[,c("iso3n", "v2x_regime", "year")], by = c("iso3n", "year"), all.x = TRUE)

# Assign NAs in developing to a 1 or 0 based on current values
count_data$developing2 = nafill(count_data$developing, type = "nocb")
count_data$developing3 = ifelse(count_data$developing2 == 0, "Developed", "Developing")

#Remove duplicate country-yr
count_data$unique_id <- paste(count_data$iso3n,count_data$year) # concatenate to make unique ID
count_data$duplicate = duplicated(count_data$unique_id) # generate the duplicate variable
count_data = count_data[count_data$duplicate != "TRUE", ] 

count_data$duplicate = duplicated(count_data$unique_id) # generate the duplicate variable

#Subset data down to 2005-2021 and non-NA values
count_data2 = count_data[count_data$year > 2004,]
count_data2 = count_data2[!is.na(count_data2$Approve),]
```

### Model 1: Change in PTA membership count
```{r}
#Build FE model on PTA count
fe_mod2 <- plm(Approve ~ Lag(total_ptas, shift = 3) + 
                 Lag(log(gdp)) + 
                 developing+
                 Lag(edu) + 
                 Lag(v2x_regime) + 
                 year,
                    data = count_data2,
                    index = c("iso3n", "year"), 
                    model = "within",
                    effect = "twoways")
summary(fe_mod2)
```

##Country Count-Level Analysis II: Change in deeper agreements

### Model 2: Change in (deeper) PTA membership count
```{r}
#obtain a list of PTAs that have a depth > mean of depth
deep_ptas = subset(desta, desta$depth_index > mean(desta$depth_index))$base_treaty


#Subset desta dyads and withdrawals based on if the base_treaty is greater than mean in desta
desta_dyads2 = desta_dyads1 %>% filter(number %in% deep_ptas)

#Count all PTAs for each country-year (to get change)
pta_count = desta_dyads2 %>% dplyr::group_by(iso3n, year) %>% dplyr::summarise(count_pta = n())

#Count all PTA withdrawals for each country-year
withdrawals2 = withdrawals %>% filter(number %in% deep_ptas)
withdrawals_count = withdrawals2 %>% dplyr::group_by(iso3n, year) %>% dplyr::summarise(count_withdrawals = n())

#Merge withdrawals column to desta_country_yr
deep_count_data = merge(pta_count, withdrawals_count, by = c("year", "iso3n"), all.x = TRUE, all.y = TRUE)

#Add missing years
deep_count_data <- setDT(deep_count_data)[CJ(iso3n=iso3n, year=seq(min(year), 2021), unique=TRUE),
              on=.(iso3n, year), roll=F]

#fill PTA count with 0
deep_count_data <- deep_count_data %>% mutate(count_pta = ifelse(is.na(count_pta), 0, count_pta),
                                                  count_withdrawals = ifelse(is.na(count_withdrawals), 0, count_withdrawals))

#Add the PTAs
deep_count_data = deep_count_data %>% group_by(iso3n) %>%
    mutate(across(count_pta, ~ accumulate(., `+`)))
deep_count_data = deep_count_data %>% group_by(iso3n) %>%
    mutate(across(count_withdrawals, ~ accumulate(., `+`)))

#Create total PTA column
deep_count_data$total_ptas = deep_count_data$count_pta - deep_count_data$count_withdrawals

## Merge approval ratings w/data ##
deep_count_data = merge(deep_count_data, gallup, by = c("iso3n", "year"), all.x = TRUE)

## Merge HDI w/data ##
deep_count_data = merge(deep_count_data, hdi_hdi, by = c("iso3n", "year"), all.x = T) %>% mutate(developing = case_when(hdi_score < 0.8 ~ 1,
                                           hdi_score>= 0.8 ~ 0))
# Assign NAs in developing to a 1 or 0 based on current values
deep_count_data$developing2 = nafill(deep_count_data$developing, type = "nocb")
deep_count_data$developing3 = ifelse(deep_count_data$developing2 == 0, "Developed", "Developing")


## Merge Control variables w/data ##
deep_count_data = merge(deep_count_data, hdi_gdp, by = c("iso3n", "year"), all.x = T)
deep_count_data = merge(deep_count_data, hdi_edu, by = c("iso3n", "year"), all.x = T)
deep_count_data = merge(deep_count_data, vdem[,c("iso3n", "v2x_regime", "year")], by = c("iso3n", "year"), all.x = TRUE)

#Remove duplicate country-yr
deep_count_data$unique_id <- paste(deep_count_data$iso3n,deep_count_data$year) # concatenate to make unique ID
deep_count_data$duplicate = duplicated(deep_count_data$unique_id) # generate the duplicate variable
deep_count_data = deep_count_data[deep_count_data$duplicate != "TRUE", ]  

#Remove NAs 
deep_count_data = deep_count_data[!is.na(deep_count_data$Approve),]

#Subset to 2005-2021 and rows without missing data
deep_count_data2 = deep_count_data[deep_count_data$year > 2004,]
deep_count_data2 = deep_count_data2[!is.na(deep_count_data2$Approve),]
```

```{r}
#Build FE model on PTA count
fe_mod3 <- plm(Approve ~ 
                 Lag((total_ptas), shift = 3) + 
                 developing +
                 Lag(log(gdp)) + 
                 Lag(edu) + 
                 Lag(v2x_regime) + 
                 year, 
                    data = deep_count_data2,
                    index = c("iso3n", "year"), 
                    model = "within",
                    effect = "twoways")
summary(fe_mod3)
```
